Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-01-30 paper.tex
neural network figure

  \node[latent] at (1,3) (d){$\mathcal{D}$};
  \node[latent] at (2.5,3) (t){$\theta$};
  \node[latent] at (1,1.5) (z){$p$};
  \node[latent] at (2.5,1.5) (tx){$\theta_x$};
  \node[obs] at (1,0) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(tx)(z)(x)}{$N$};
  \draw [->,dashed] (x.east) to[out = 0,in = -90] node[fill = white]{$q(x)$} (tx.south);
  \draw [->,dashed] (d.west) to[out = 180,in = 180] (z.west);
  \draw [->,dashed] (tx.west) -- (z.east);
  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe program outputs $x$. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta_x = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

@@

\begin{tabular}{ll}
  \toprule
  Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
  & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
  \midrule
  Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
  \\\midrule
  \begin{tabular}{l}
    Subset of \\Learned DSL
    \end{tabular}&
  \begin{tabular}{ll}
    $\lambda x. \mathcal{R}\times x + \mathcal{R}$ &linear\\
    $\lambda x. \mathcal{R} + x$ &increment\\
    $\lambda x. x\times (\text{linear } x)$&$\text{quadratic}_0$\\
    $\lambda x. \text{increment }(\text{quadratic}_0\; x)$&quadratic\\
  \end{tabular}

  \\\bottomrule
  \end{tabular}

>>>>

  \node[latent] at (1,3) (d){$\mathcal{D}$};
  \node[latent] at (2.5,3) (t){$\theta$};
  \node[latent] at (1,1.5) (z){$p$};
  \node[latent] at (3.5,1.5) (tx){$\theta_x$};
  \node[obs] at (1,-0.5) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(tx)(z)(x)}{$x\in X$};
  \draw [->,dashed] (x.east) to[out = 0,in = -90] node(nn)[fill = white]{
          \includegraphics[width = 1.5cm]{figures/MLP.png} 
  } (tx.south);
  \node at (nn.south) {$q(x)$};
  \draw [->,dashed] (d.west) to[out = 180,in = 180] (z.west);
  \draw [->,dashed] (tx.west) -- (z.east);
  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe program outputs $x$. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta_x = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

\section{The \system Algorithm}

@@

What DSL does \system learn?
The learned DSL contains templates for quadratic and linear functions,
which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain.
Examining the programs themselves,
one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
for example, it represents the polynomial $8x^2+8x$

\begin{tabular}{ll}
  \toprule
  Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
  & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
  \midrule
  Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
  \\\midrule
  \begin{tabular}{l}
    Subset of \\Learned DSL
    \end{tabular}&
  \begin{tabular}{ll}
    $\lambda x. \mathcal{R}\times x + \mathcal{R}$ &linear\\
    $\lambda x. \mathcal{R} + x$ &increment\\
    $\lambda x. x\times (\text{linear } x)$&$\text{quadratic}_0$\\
    $\lambda x. \text{increment }(\text{quadratic}_0\; x)$&quadratic\\
  \end{tabular}
  \\\bottomrule
  \end{tabular}
